Two applications of logic programming to Coq

### 2020-11-05 papers/types2020/sec3.tex
Some thinking about a good introduction to Section 3... Commiting to resolve a merge.
Minor updates.  Will return tomorrow morning.
Fix typos and a bit more text on implementation

\subsection{Proof systems with certificates}
\label{ssec:focused}
\todo{Change title? }
\todo{Use same names in Fig and 3: A for Atom, L for ListArgs etc.}

@@

%

@@

Coq terms are accessed through the Coq-Elpi API, and their representation
in \lP takes advantage of the native \lP constructs such as lists and binders.
The following is a list of some of signatures of the notions that we use:
\begin{lstlisting}
  kind term   type.   % reification of Coq terms
  kind gref    type     % reification of refernces to global terms
  type global gref -> term.
  type indt   inductive -> gref.  % reification of inductive types
  type indc   constructor -> gref. % reification of their constructors 
  type app    list term -> term.  % reification of nary application
  type prod   name -> term -> (term -> term) -> term.
\end{lstlisting}

Note that \lsti{prod} encodes dependent products taking a name for pretty printing,  a term and a \lP function from
terms to terms, namely $\forall x : A, B$ is encoded by \texttt{prod ``x'' A (x\ B x)};
when, in the implementation of the product-left rule, we
apply $B$ to the variable \lsti{Tm},
we get a term that can be used to continue backchaining. This
application is obtained via meta-level substitution, in the style of HOAS. In this sense, our calculus uses \emph{implicit} substitutions, rather than explicit ones as in LJT and PTSC's tradition, the more since proof search in our application is cut-free. 
% Syntax: one can think as if something of type \lsti{term -> term} were a
% \lsti{term} with a hole in it for another \lsti{term}.
\todo{Reput safe-dest-app in code, a bit more details about the unfold case}
The unfolding rule makes use of the Coq-Elpi primitives
\lsti{coq.safe-dest-app} in order to obtain the head term of a (possibly
nested) application, and \lsti{coq.env.indt} in order to access the global
environment of inductive definitions and query for information about them.

\smallskip
In figure~\ref{fig:augmented}  each
inference rule is augmented with an additional premise involving an
\emph{expert predicate}, a certificate $\Xi$, and possibly continuations of
certificates ($\Xi'$, $\Xi_1$, $\Xi_2$) with extracted information from
These notions are taken from the general setting of \emph{foundational
  proof certificates}~\cite{chihani17jar}.
%
In our case here, an FPC is a collection of \lP clauses that
provide the remaining details not supplied in Figure~\ref{fig:kernel}:
that is, the exact set of constructors for the  type of certificates as
well as the specification of the expert predicates listed \emph{ibidem}.
%
The top of Figure~\ref{fig:resources} displays two such FPCs,
both of which can be used to describe proofs where we bound
the dimension of  a proof.
%
For example, the first FPC dictates that the query \mbox{\lsti{(check (qheight 5) B)}} is
provable in  the kernel using the clauses in Figures~\ref{fig:kernel}
and~\ref{fig:resources} if and only if the height of that proof is 5
or less.
%
Similarly, the second FPC can be used to bound the total number of instances of
unfoldings in a proof.
%
%In particular, the query \mbox{\lsti{(check (qsize 5 H) B)}}
% \begin{lstlisting}
% sigma H\ (check (qsize 5 H) B)
% \end{lstlisting}
%is provable if and only if the total number of unfoldings of that
%proof is 5 or less.

>>>>

\subsection{Proof systems with certificates}
\label{ssec:focused}
\todo{Change title? }

@@

@@

Coq terms are accessed through the Coq-Elpi API, and their representation
in \lP takes advantage of the native \lP constructs such as lists and binders.
The following is a list of some of the signatures of the notions that we use:
\begin{lstlisting}
  kind term   type.   % reification of Coq terms
  kind gref   type     % reification of refernces to global terms
  type global gref -> term.
  type indt   inductive -> gref.  % reification of inductive types
  type indc   constructor -> gref. % reification of their constructors 
  type app    list term -> term.  % reification of nary application
  type prod   name -> term -> (term -> term) -> term.
\end{lstlisting}

Note that \lsti{prod} encodes dependent products taking a name for pretty
printing, a term and a \lP function from terms to terms, namely $\forall x :
A, B$ is encoded by \texttt{prod ``x'' A (x\textbackslash{} B x)}; when, in the
implementation of the product-left rule, we apply $B$ to the variable
\lsti{Tm}, we get a term that can be used to continue backchaining. This
application is obtained via meta-level substitution, in the style of HOAS. In
this sense, our calculus uses \emph{implicit} substitutions, rather than
explicit ones as in LJT and PTSC's tradition, the more since proof search in
our application is cut-free.
% Syntax: one can think as if something of type \lsti{term -> term} were a
% \lsti{term} with a hole in it for another \lsti{term}.
The unfolding rule makes use of the Coq-Elpi primitives
\lsti{coq.safe-dest-app} in order to obtain the head term of a (possibly
nested) application, and \lsti{coq.env.indt} in order to access the global
environment of inductive definitions and query for information about them.
The \lsti{unfold_expert} predicate, then, has the role of selecting which
constructor to introduce for the inductive type. The kernel will successively
obtain the type of the selected constructor, and initiate the backchaining
phase on it.

\smallskip
In figure~\ref{fig:augmented}  each
inference rule is augmented with an additional premise involving an
\emph{expert predicate}, a certificate $\Xi$, and possibly continuations of
certificates ($\Xi'$, $\Xi_1$, $\Xi_2$) with extracted information from
These notions are taken from the general setting of \emph{foundational
  proof certificates}~\cite{chihani17jar}.
%
\todo{MM: Talk about randomized FPC?}
In our case here, an FPC is a collection of \lP clauses that
provide the remaining details not supplied in Figure~\ref{fig:kernel}:
that is, the exact set of constructors for the  type of certificates as
well as the specification of the expert predicates listed \emph{ibidem}.
%
The top of Figure~\ref{fig:resources} displays two such FPCs,
both of which can be used to describe proofs where we bound
the dimension of  a proof.
%
For example, the first FPC dictates that the query \mbox{\lsti{(check (qheight 5) B)}} is
provable in  the kernel using the clauses in Figures~\ref{fig:kernel}
and~\ref{fig:resources} if and only if the height of that proof is 5
or less.
%
Similarly, the second FPC can be used to bound the total number of instances of
unfoldings in a proof.
%
%In particular, the query \mbox{\lsti{(check (qsize 5 H) B)}}
% \begin{lstlisting}
% sigma H\ (check (qsize 5 H) B)
% \end{lstlisting}
%is provable if and only if the total number of unfoldings of that
%proof is 5 or less.
